---
title: "Turnout Data Analysis"
author: "Daniel Marcelino"
date: "October 10, 2013"
output: ioslides_presentation
---


## Using a subset of 2000 US election data

- Analyzing voter turnout from the 2000 election
- Overall voter turnout 51.21%
- Why so low?
- barriers, disenfranchisement, Tuesday voting, optional
- Discrepencies in turnout
- Implications

## Understanding the data
- Source: Census Bureau Current Population Survey sample
1500 observations
- vote: 1 means vote , 0 means did not (mean=.8553)
- income: 1-17 where 1 is <$5,000 and 17 is > $75,000
- education: 1-4 where 1 represents some high school and 4 represents more than college
- age: 18-85
- female: 1= female, 0=male
- state: They only looked at AR and SC


## How does this data compare with national data?

AR and SC, both rural Southern states
SC 28.5% black, AR 15.6% black, US 12% black
Per capita income: SC $42k, AR $38k, US $48k
Median age: SC 36.9 , AR 36.6, US 36.5
2000 voter turnout: SC 46.6 %, AR 47.8%, US 51.2%

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

## Required packages and data

```{r}
library(Zelig)
library(stargazer)
library(dplyr)
library(ggplot2)
library(SciencesPo)
data(voteincome)
summary(voteincome)
```


## Some demographics 
Men are 44% of this study, women are 56% 

```{r}
hist(voteincome$income)
hist(voteincome$age)
hist(voteincome$education)
```


## Linear Models
```{r}
m1<- zelig(vote~income, model="ls", data=voteincome)
m2<- zelig(vote~education, model="ls", data=voteincome)
m3<- zelig(vote~age, model="ls", data=voteincome)
m4<- zelig(vote~female, model="ls", data=voteincome)
m5<- zelig(vote~income+age, model="ls", data=voteincome)
m6<- zelig(vote~income+education+age+female, model="ls", data=voteincome)
```


## Table for comparing the models
```{r}
stargazer(m1, m2, m3, m4, m5, m6, type="html", title="Linear Regression Models")
```


## Main evidence from the models
- Singificant Variables
- When alone: income, education, and age
- In m6, everything is significant -R^2
- Everything is low, highest is .037. Meaning, there is something else there
- Now let's graph our regressions to better visualize them



## Make exploratory plots of the linear evidence
```{r}
g <- ggplot(voteincome, mapping=aes(x=income, y=vote))
g1 <- g+ geom_point() + stat_smooth(method="lm")
g1
```

## Make exploratory plots of the linear evidence
```{r}
g3 <- ggplot(voteincome, mapping=aes(x=education, y=vote))+ geom_point() + stat_smooth(method="lm")
g3
```

## Make exploratory plots of the linear evidence
```{r}
g4 <- ggplot(voteincome, mapping=aes(x=age, y=vote))+ geom_point() + stat_smooth(method="lm")
g4
```

## Make exploratory plots of the linear evidence
```{r}
g5 <- ggplot(voteincome, mapping=aes(x=female, y=vote))+ geom_point() + stat_smooth(method="lm")
g5
```

## Make exploratory plots of the linear evidence
```{r}
qplot(income, age, data=voteincome, colour=factor(vote))
```


```{r}
x1 <- setx(m6, female= 1)
s1 <- sim(m6, x=x1)
summary(s1)
```


```{r}
x1 <- setx(m6, female= 0)
s2 <- sim(m6, x=x1)
summary(s2)
```



```{r}
inc<-min(voteincome$income):max(voteincome$income)
x2 <- setx(m6, female=1, income=inc)
s3 <-sim (m6, x=x2)
summary(s3)
```


```{r}
plot.ci(s3)
```


```{r}
inc<-min(voteincome$income):max(voteincome$income)
x3 <- setx(m6, female=0, income=inc)
s4 <-sim (m6, x=x3)

summary(s4)
```

## Inspection of the confidence intervals 
```{r}
plot.ci(s4)
```

## Final remarks
- could be Very different from nat'l data
- this data does not represent SC and AR
- income, education, and age (not gender)
- total turnout 85.5% -more likely to turnout: wealthy, educated, older, women
