---
title: "Olympic 100m Men Final Predictions"
author: "Daniel Marcelino"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

http://sportsdemo1.gracenote.com/documentation/vmt#demo

https://www.iaaf.org/records/by-discipline/sprints/100-metres/outdoor/men


Usain Bolt 'will not win Rio 2016 100m final'. So who will?

suggests the United States will once again lead the way, ahead of China, while Britain suffers a spectacular slump -- and Bolt misses out on men's 100m gold for the first time since 2004.

Jamaican legend Bolt, 28, is predicted to miss out on Olympic 100m gold next year.
"We've got Bolt second at the moment, behind Justin Gatlin," says Gleave.
Gatlin, the American who served a four-year drugs ban between 2006 and 2010, won bronze at London 2012 -- where Bolt completed his second successive Olympic treble, winning 100m, 200m and 4x100m relay gold meals.
Bolt has dominated the 100m for the better part of a decade, winning the world titles in 2009 and 2013 alongside his two Olympic gold medals. Only in 2011, when he false-started, has he missed out.

```{r}
df <- data.frame(
"Year", "Official results"
1908	10.8
1912	10.8
1920	10.8
1924	10.6
1928	10.8
1932	10.3
1936	10.3
1948	10.3
1952	10.4
1956	10.5
1960	10.2
1964	10
1968	9.95
1972	10.14
1976	10.06
1980	10.25
1984	9.99
1988	9.92
1992	9.96
1996	9.84
2000	9.87
2004	9.85
2008	9.69
2012	9.63
2016	
2020	
)
```

